---
title: "Predictive Modeling Project"
author: "Shinya Hashimoto"
date: "2024-02-23"
output: 
  html_document:
    keep_md: true
---

### Overview
This project involves constructing a machine learning algorithm to predict the quality of activity from an activity monitor using data collected from accelerometers on wearable devices. The primary goal is to predict the manner (`classe` variable) in which exercises were performed.

## Repository Contents

- This document in both `.Rmd` (source) and `.html` (compiled) formats.
- Additional scripts and datasets used in the analysis.

### Data Loading and Preprocessing

```{r, cache=TRUE}
library(caret)
library(ggplot2)
library(rpart.plot)
library(corrplot)
set.seed(12345)

# Data download URLs
trainingUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testingUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

# Reading data
training <- read.csv(url(trainingUrl), na.strings=c("NA","#DIV/0!", ""))
testing <- read.csv(url(testingUrl), na.strings=c("NA","#DIV/0!", ""))
```

### Data Cleaning and Preprocessing

```{r, cache=TRUE}
# Remove columns with NA values
training <- training[, colSums(is.na(training)) == 0]
testing <- testing[, colSums(is.na(testing)) == 0]

# Remove the first 7 columns
training <- training[, -c(1:7)]
testing <- testing[, -c(1:7)]
```

### Model Building and Evaluation

```{r, cache=TRUE}
subSamples <- createDataPartition(y=training$classe, p=0.75, list=FALSE)
subTraining <- training[subSamples,]
subTesting <- training[-subSamples,]

# Setting training control
control <- trainControl(method="cv", number=3, verboseIter=TRUE)

# Building and evaluating models
models <- c("rpart", "rf", "gbm", "svmLinear")
names(models) <- c("Decision Tree", "Random Forest", "Gradient Boosted Trees", "SVM")

results <- lapply(models, function(method) {
  # Build model
  model <- train(classe ~ ., data=subTraining, method=method, trControl=control, tuneLength=1)
  # Predict on test data
  prediction <- predict(model, subTesting)
  # Evaluate predictions
  cm <- confusionMatrix(prediction, factor(subTesting$classe))
  # Return confusion matrix and model
  return(list(cm = cm, model = model))
})

# Display results (accuracy only)
results_df <- sapply(results, function(x) x$cm$overall['Accuracy'])
data.frame(Model=names(models), Accuracy=unlist(results_df), row.names=NULL)
```

## Predictions on Test Set
Conduct test data estimation with the best performing model.

```{r, cache=TRUE}
# Predictions on test set
# Using Random Forest model for predictions
# Retrieving model based on 'rf' key
best_model <- results[["Random Forest"]]$model

# Execute prediction on test dataset
pred <- predict(best_model, subTesting)

# Display prediction results
print(pred)
```
